---
title: "AN INVESTIGATION OF FEATURE SELECTION METHODS FOR CLASSIFICATION PURPOSES USING SUPERVISED MACHINE LEARNING"
author: "M.Sc. Buse Nur BALTACIOĞLU ~ Assoc. Prof. İdil YAVUZ"
date: "2022-12-02"
output:
  html_document:
    toc: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Data

<https://www.kaggle.com/datasets/yasserhessein/early-stage-diabetes-risk-prediction-dataset>

<https://archive-beta.ics.uci.edu/dataset/529/early+stage+diabetes+risk+prediction+dataset>

Age 1.20-65 (y<="mean=48"<n)
<br>
Gender 1. Male, 2.Female
<br>
Polyuria 1.Yes, 2.No
<br>
Polydipsia 1.Yes, 2.No
<br>
sudden weight loss 1.Yes, 2.No
<br>
weakness 1.Yes, 2.No
<br>
Polyphagia 1.Yes, 2.No
<br>
Genital thrush 1.Yes, 2.No
<br>
visual blurring 1.Yes, 2.No
<br>
Itching 1.Yes, 2.No
<br>
Irritability 1.Yes, 2.No
<br>
delayed healing 1.Yes, 2.No
<br>
partial paresis 1.Yes, 2.No
<br>
muscle stiness 1.Yes, 2.No
<br>
Alopecia 1.Yes, 2.No
<br>
Obesity 1.Yes, 2.No
<br>
Class 1.Positive, 2.Negative
<br>

```{r message=FALSE, warning=FALSE}
library(readr)
df <- read_csv("diabetes_data_upload.csv")
```

```{r}
summary(df$Age)
```

```{r}
df$Age=as.factor(ifelse(df$Age <= 48, "0", "1"))
df$Gender=as.factor(df$Gender)
df$Polyuria=as.factor(df$Polyuria)
df$Polydipsia=as.factor(df$Polydipsia)
df$`sudden weight loss`=as.factor(df$`sudden weight loss`)
df$weakness=as.factor(df$weakness)
df$Polyphagia=as.factor(df$Polyphagia)
df$`Genital thrush`=as.factor(df$`Genital thrush`)
df$`visual blurring`=as.factor(df$`visual blurring`)
df$Itching=as.factor(df$Itching)
df$Irritability=as.factor(df$Irritability)
df$`delayed healing`=as.factor(df$`delayed healing`)
df$`partial paresis`=as.factor(df$`partial paresis`)
df$`muscle stiffness`=as.factor(df$`muscle stiffness`)
df$Alopecia=as.factor(df$Alopecia)
df$Obesity=as.factor(df$Obesity)
df$class=as.factor(df$class)
```

```{r}
colnames(df)=c("Age","Gender","Polyuria","Polydipsia","Sudden.weight.loss",
                 "weakness","Polyphagia","Genital.thrush","visual.blurring",
                 "Itching","Irritability","delayed.healing","partial.paresis",
                 "muscle.stiffness","Alopecia","Obesity","Class")
```

```{r}
summary(df)
```

# Full Model Glm

```{r message=FALSE, warning=FALSE}
library(caret)
```

```{r}
set.seed(8)
Train=createDataPartition(df$Class,
                          p=0.96,
                          list = FALSE)
training=df[Train,]
testing=df[-Train,]
dim(training)
dim(testing)
```

```{r}
summary(testing)
```

```{r}
summary(training)
```

```{r}
set.seed(8)
ctrl=trainControl(method = "cv",
                      number = 10,
                      savePredictions=TRUE,
                      classProbs=TRUE)
```

```{r}
set.seed(8)
model=train(Class~.,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model)
```

```{r}
summary(model)
```

```{r}
pred_train<-predict(model,
                    newdata = training, 
                    type = "raw")
pred_test<-predict(model,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_train, training$Class)
```

```{r}
table(pred_test, testing$Class)
```

```{r include=FALSE}
(sens_train=yardstick::sens_vec(training$Class, pred_train))
(sens_test=yardstick::sens_vec(testing$Class, pred_test))
```

```{r include=FALSE}
(spec_train=yardstick::specificity_vec(training$Class, pred_train))
(spec_test=yardstick::specificity_vec(testing$Class, pred_test))
```

```{r}
library(ROCR)
```

```{r}
prob=predict(model,
             newdata=testing, 
             type="raw")
pred=prediction(as.numeric(prob),
                testing$Class)
perf=performance(pred,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf)
```

```{r}
auc=performance(pred, measure = "auc")
(auc=auc@y.values[[1]])
```

# Number of Feature = 6

## Information gain

```{r}
library(FSelector)
```

```{r}
weights_inf <- information.gain(Class~., training)

subset_inf <- cutoff.k(weights_inf, 6)

(f_inf <- as.simple.formula(subset_inf, "Class"))
```

```{r}
set.seed(8)
model_inf=train(Class ~ Polyuria + Polydipsia + Gender + Sudden.weight.loss + 
    partial.paresis + Polyphagia,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model_inf)
```

```{r}
summary(model_inf)
```

```{r}
pred_inf_train<-predict(model_inf,
                    newdata = training, 
                    type = "raw")
pred_inf_test<-predict(model_inf,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_inf_train, training$Class)
```

```{r}
table(pred_inf_test, testing$Class)
```

```{r include=FALSE}
(sens_inf_train=yardstick::sens_vec(training$Class, pred_inf_train))
(sens_inf_test=yardstick::sens_vec(testing$Class, pred_inf_test))
```

```{r include=FALSE}
(spec_inf_train=yardstick::specificity_vec(training$Class, pred_inf_train))
(spec_inf_test=yardstick::specificity_vec(testing$Class, pred_inf_test))
```

```{r}
prob_inf=predict(model_inf,
             newdata=testing, 
             type="raw")
pred_inf=prediction(as.numeric(prob_inf),
                testing$Class)
perf_inf=performance(pred_inf,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf_inf)
```

```{r}
auc_inf=performance(pred_inf, measure = "auc")
(auc_inf=auc_inf@y.values[[1]])
```

## Gain ratio

```{r}
weights_gain <- gain.ratio(Class~., training)

subset_gain <- cutoff.k(weights_gain, 6)

(f_gain <- as.simple.formula(subset_gain, "Class"))
```

```{r}
set.seed(8)
model_gain=train(Class ~ Polyuria + Polydipsia + Gender + Sudden.weight.loss + 
    partial.paresis + Irritability,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model_gain)
```

```{r}
summary(model_gain)
```

```{r}
pred_gain_train<-predict(model_gain,
                    newdata = training, 
                    type = "raw")
pred_gain_test<-predict(model_gain,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_gain_train, training$Class)
```

```{r}
table(pred_gain_test, testing$Class)
```

```{r include=FALSE}
(sens_gain_train=yardstick::sens_vec(training$Class, pred_gain_train))
(sens_gain_test=yardstick::sens_vec(testing$Class, pred_gain_test))
```

```{r include=FALSE}
(spec_gain_train=yardstick::specificity_vec(training$Class, pred_gain_train))
(spec_gain_test=yardstick::specificity_vec(testing$Class, pred_gain_test))
```

```{r}
prob_gain=predict(model_gain,
             newdata=testing, 
             type="raw")
pred_gain=prediction(as.numeric(prob_gain),
                testing$Class)
perf_gain=performance(pred_gain,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf_gain)
```

```{r}
auc_gain=performance(pred_gain, measure = "auc")
(auc_gain=auc_gain@y.values[[1]])
```

## Symmetrical uncertainty

```{r}
weights_sym <- symmetrical.uncertainty(Class~., training)

subset_sym <- cutoff.k(weights_sym, 6)

(f_sym <- as.simple.formula(subset_sym, "Class"))
```

```{r}
set.seed(8)
model_sym=train(Class ~ Polyuria + Polydipsia + Gender + Sudden.weight.loss + 
    partial.paresis + Polyphagia,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model_sym)
```

```{r}
summary(model_sym)
```

```{r}
pred_sym_train<-predict(model_sym,
                    newdata = training, 
                    type = "raw")
pred_sym_test<-predict(model_sym,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_sym_train, training$Class)
```

```{r}
table(pred_sym_test, testing$Class)
```

```{r include=FALSE}
(sens_sym_train=yardstick::sens_vec(training$Class, pred_sym_train))
(sens_sym_test=yardstick::sens_vec(testing$Class, pred_sym_test))
```

```{r include=FALSE}
(spec_sym_train=yardstick::specificity_vec(training$Class, pred_sym_train))
(spec_sym_test=yardstick::specificity_vec(testing$Class, pred_sym_test))
```

```{r}
prob_sym=predict(model_sym,
             newdata=testing, 
             type="raw")
pred_sym=prediction(as.numeric(prob_sym),
                testing$Class)
perf_sym=performance(pred_sym,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf_sym)
```

```{r}
auc_sym=performance(pred_sym, measure = "auc")
(auc_sym=auc_sym@y.values[[1]])
```

## Chi-square

```{r}
weights_chi <- chi.squared(Class~., training)

subset_chi <- cutoff.k(weights_chi, 6)

(f_chi <- as.simple.formula(subset_chi, "Class"))
```

```{r}
set.seed(8)
model_chi=train(Class ~ Polyuria + Polydipsia + Gender + Sudden.weight.loss + 
    partial.paresis + Polyphagia,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model_chi)
```

```{r}
summary(model_chi)
```

```{r}
pred_chi_train<-predict(model_chi,
                    newdata = training, 
                    type = "raw")
pred_chi_test<-predict(model_chi,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_chi_train, training$Class)
```

```{r}
table(pred_chi_test, testing$Class)
```

```{r include=FALSE}
(sens_chi_train=yardstick::sens_vec(training$Class, pred_chi_train))
(sens_chi_test=yardstick::sens_vec(testing$Class, pred_chi_test))
```

```{r include=FALSE}
(spec_chi_train=yardstick::specificity_vec(training$Class, pred_chi_train))
(spec_chi_test=yardstick::specificity_vec(testing$Class, pred_chi_test))
```

```{r}
prob_chi=predict(model_chi,
             newdata=testing, 
             type="raw")
pred_chi=prediction(as.numeric(prob_chi),
                testing$Class)
perf_chi=performance(pred_chi,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf_chi)
```

```{r}
auc_chi=performance(pred_chi, measure = "auc")
(auc_chi=auc_chi@y.values[[1]])
```

## Cramer

```{r message=FALSE, warning=FALSE}
library(FSinR)
```

```{r}
cramer_evaluator <- cramer()
weights_cra<-cramer_evaluator(training,'Class',c("Age","Gender","Polyuria","Polydipsia","Sudden.weight.loss",
                 "weakness","Polyphagia","Genital.thrush","visual.blurring",
                 "Itching","Irritability","delayed.healing","partial.paresis",
                 "muscle.stiffness","Alopecia","Obesity"))

sort(weights_cra, decreasing = TRUE)
```

```{r}
set.seed(8)
model_cra=train(Class ~ Polyuria + Polydipsia + Gender + Sudden.weight.loss + 
    partial.paresis + Polyphagia,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model_cra)
```

```{r}
summary(model_cra)
```

```{r}
pred_cra_train<-predict(model_cra,
                    newdata = training, 
                    type = "raw")
pred_cra_test<-predict(model_cra,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_cra_train, training$Class)
```

```{r}
table(pred_cra_test, testing$Class)
```

```{r include=FALSE}
(sens_cra_train=yardstick::sens_vec(training$Class, pred_cra_train))
(sens_cra_test=yardstick::sens_vec(testing$Class, pred_cra_test))
```

```{r include=FALSE}
(spec_cra_train=yardstick::specificity_vec(training$Class, pred_cra_train))
(spec_cra_test=yardstick::specificity_vec(testing$Class, pred_cra_test))
```

```{r}
prob_cra=predict(model_cra,
             newdata=testing, 
             type="raw")
pred_cra=prediction(as.numeric(prob_cra),
                testing$Class)
perf_cra=performance(pred_cra,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf_cra)
```

```{r}
auc_cra=performance(pred_cra, measure = "auc")
(auc_cra=auc_cra@y.values[[1]])
```

## OneR

```{r}
weights_oneR <- oneR(Class~., training)

subset_oneR <- cutoff.k(weights_oneR, 6)

(f_oneR <- as.simple.formula(subset_oneR, "Class"))
```

```{r}
set.seed(8)
model_oneR=train(Class ~ Polyuria + Polydipsia + Sudden.weight.loss + partial.paresis + 
    Gender + Alopecia,
            data = training,
            method="glm",
            family="binomial",
            trControl=ctrl)
```

```{r}
print(model_oneR)
```

```{r}
summary(model_oneR)
```

```{r}
pred_oneR_train<-predict(model_oneR,
                    newdata = training, 
                    type = "raw")
pred_oneR_test<-predict(model_oneR,
                   newdata = testing,
                   type = "raw")
```

```{r}
table(pred_oneR_train, training$Class)
```

```{r}
table(pred_oneR_test, testing$Class)
```

```{r include=FALSE}
(sens_oneR_train=yardstick::sens_vec(training$Class, pred_oneR_train))
(sens_oneR_test=yardstick::sens_vec(testing$Class, pred_oneR_test))
```

```{r include=FALSE}
(spec_oneR_train=yardstick::specificity_vec(training$Class, pred_oneR_train))
(spec_oneR_test=yardstick::specificity_vec(testing$Class, pred_oneR_test))
```

```{r}
prob_oneR=predict(model_oneR,
             newdata=testing, 
             type="raw")
pred_oneR=prediction(as.numeric(prob_oneR),
                testing$Class)
perf_oneR=performance(pred_oneR,
                 measure = "tpr",
                 x.measure = "fpr")
plot(perf_oneR)


```

```{r}
auc_oneR=performance(pred_oneR, measure = "auc")
(auc_oneR=auc_oneR@y.values[[1]])
```

# Result

```{r include=FALSE}
Cv_acc_inf=model_inf$results[2]
Cv_kappa_inf=model_inf$results[3]
Cv_aic_inf=model_inf$finalModel$aic
```

```{r include=FALSE}
Cv_acc_gain=model_gain$results[2]
Cv_kappa_gain=model_gain$results[3]
Cv_aic_gain=model_gain$finalModel$aic
```

```{r include=FALSE}
Cv_acc_sym=model_sym$results[2]		
Cv_kappa_sym=model_sym$results[3]
Cv_aic_sym=model_sym$finalModel$aic
```

```{r include=FALSE}
Cv_acc_chi=model_chi$results[2]	
Cv_kappa_chi=model_chi$results[3]
Cv_aic_chi=model_chi$finalModel$aic
```

```{r include=FALSE}
Cv_acc_cra=model_cra$results[2]		
Cv_kappa_cra=model_cra$results[3]
Cv_aic_cra=model_cra$finalModel$aic
```

```{r include=FALSE}
Cv_acc_oneR=model_oneR$results[2]		
Cv_kappa_oneR=model_oneR$results[3]
Cv_aic_oneR=model_oneR$finalModel$aic
```

```{r}
Cv_acc=model$results[2]
Cv_kappa=model$results[3]
Cv_aic=model$finalModel$aic
```

```{r include=FALSE}
r_acc=rbind(auc, auc_inf, auc_gain, auc_sym, auc_chi, auc_cra, auc_oneR)

r_cv_acc=rbind(Cv_acc, Cv_acc_inf, Cv_acc_gain, Cv_acc_sym, Cv_acc_chi, Cv_acc_cra, Cv_acc_oneR)

r_cv_kappa=rbind(Cv_kappa,Cv_kappa_inf, Cv_kappa_gain, Cv_kappa_sym, Cv_kappa_chi, Cv_kappa_cra, Cv_kappa_oneR)

r_cv_aic=rbind(Cv_aic,Cv_aic_inf, Cv_aic_gain, Cv_aic_sym, Cv_aic_chi, Cv_aic_cra, Cv_aic_oneR)

```

```{r include=FALSE}
r=data.frame(r_acc,r_cv_acc,r_cv_kappa,r_cv_aic)

row.names(r)=c("Full Model","Information Gain","Gain Ratio","Symmetrical Uncertainty","Chi-square","Cramer","OneR")
colnames(r)=c("Test Accurancy", "CV Accurancy", "CV Kappa", "CV AIC")
```

```{r}
d=data.frame(r_acc,r_cv_acc)
row.names(d)=c("Full Model","Information Gain","Gain Ratio","Symmetrical Uncertainty","Chi-square","Cramer","OneR")
colnames(d)=c("Validation Accurancy", "Cross Validation Accurancy")
```

```{r}
rmarkdown::paged_table(d)
```

```{r}
plot(perf_inf, col="deepskyblue", lwd=2,
     main="Validation Accurancy")
plot(perf_oneR, col="coral",lwd=2, add=T)
abline(0,1)
```

```{r include=FALSE}
specc=rbind(spec_train,spec_inf_train,spec_gain_train,spec_sym_train,spec_chi_train,spec_cra_train,spec_oneR_train)
senss=rbind(sens_train,sens_inf_train,sens_gain_train,sens_sym_train,sens_chi_train,sens_cra_train,sens_oneR_train)
c_train=data.frame(senss,specc)

row.names(c_train)=c("Full Model","Information Gain","Gain Ratio","Symmetrical Uncertainty","Chi-square","Cramer","OneR")
colnames(c_train)=c("Sensitivity","Specificity")
```

```{r include=FALSE}
spec=rbind(spec_test,spec_inf_test,spec_gain_test,spec_sym_test,spec_chi_test,spec_cra_test,spec_oneR_test)
sens=rbind(sens_test,sens_inf_test,sens_gain_test,sens_sym_test,sens_chi_test,sens_cra_test,sens_oneR_test)
c_test=data.frame(sens,spec)

row.names(c_test)=c("Full Model","Information Gain","Gain Ratio","Symmetrical Uncertainty","Chi-square","Cramer","OneR")
colnames(c_test)=c("Sensitivity","Specificity")
```

```{r}
rmarkdown::paged_table(c_train)
```

```{r}
rmarkdown::paged_table(c_test)
```

```{r include=FALSE}
gain_ratio=c(0.75,0.8700736,0.8939936,0.8939936,0.8939936,0.8639904,0.8639904,
             0.8939528,0.8940328,0.8900328,0.8900328,0.9059936,0.9099936,
             0.9099936,0.9119544,0.9583333)
oner=c(0.75,0.8700736,0.8700736,0.8721144,0.8939936,0.9059952,0.9039544,
       0.9039544,0.8999920,0.9059152,0.9059152,0.9098743,0.9180360,
       0.9180360,0.9140360,0.9583333)
symmetric=c(0.75,0.8700736,0.8939936,0.8939936,0.8939936,0.8919936,0.8639904,
            0.8939528,0.8940328,0.8900328,0.8900328,0.9059936,0.9099936,
            0.9099936,0.9119544,0.9583333)
other=c(0.75,0.8700736,0.8939936,0.8939936,0.8939936,0.8919936,0.8639904,
        0.8939528,0.8940328,0.8900328,0.8900328,0.9059936,0.9079936,
        0.9099936,0.9119544,0.9583333)
```

```{r include=FALSE}
c=c(gain_ratio,symmetric,oner,other)
method=c("gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio",
    "gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio","gain_ratio",
    "symmetric","symmetric","symmetric","symmetric","symmetric","symmetric","symmetric","symmetric","symmetric",
    "symmetric","symmetric","symmetric","symmetric","symmetric","symmetric","symmetric",
    "oner","oner","oner","oner","oner","oner","oner",
    "oner","oner","oner","oner","oner","oner","oner","oner","oner",
    "other","other","other","other","other","other","other","other","other",
    "other","other","other","other","other","other","other")
f=c(1:16,1:16,1:16,1:16)
cvaa=data.frame(f,method,c)
```

```{r}
library(ggplot2)
```

```{r}
ggplot(cvaa, aes(x=f, y=c, color=method))+
  geom_point(aes(color = method, shape=method))+
  scale_color_brewer(palette="Dark2")+
  geom_line(aes(color = method))+
  labs(title = "Accuracy plot from cross validation by number of features", 
       x = "Number of features", y="CV accurancy")+
  theme(plot.title = element_text(color = "#828282", size = 12, face = "bold"),
        axis.title.x = element_text(color = "darkgrey", size = 10, face = "bold"),
        axis.title.y = element_text(color = "darkgrey", size = 10, face = "bold"),
        legend.title = element_text(color = "#828282", size = 10, face = "bold"),
        legend.background = element_rect(fill = "lightyellow"))
```














